Solving the maximum independant set problem by generic programming einsum networks

### 2021-06-16 paper/paper.tex
save

One may also want to obtain all ground state configurations, it can be achieved replacing the factors $a_x$ with a set of bit strings $s_x$.
We design a new element type that having algebra
\iffalse
\begin{align}
\begin{split}
    s_x \oplus s_y &= s_x \cup s_y\\
    s_x \odot s_y &= \{\sigma \lor \tau | \sigma \in s_x, \tau \in s_y\}\\
    \mymathbb{0} &= \{\}\\
    \mymathbb{1} &= \{\boldsymbol 0\}
\end{split}
\end{align}
\fi
\begin{align}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases},\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
    \mymathbb{0} &= \{\}\infty^{-\infty},\\
    \mymathbb{1} &= \{\boldsymbol 0\}\infty^{0},
\end{split}
\end{align}
where $\vee$ is the bit-wise \texttt{or}. One can easily check that this replacement does not change the fact that the algebra is a commutative semiring.
We first initialize the bit strings of the variable $x$ in the vertex tensor to a vertex index $i$ dependent onehot vector $x_i = \boldsymbol{e}_{i}$,
then we contract the tensor network. The resulting object will give us the set of all optimal configurations.
By slightly modifying the above algebra, it can also be used to obtain just a single configuration to save computational effort.
We leave this as an exercise for readers.

@@

Moving $B_{ik}$ to the right hand side, we have
\begin{align}
    A_{ij} &\leq (\oplus_{k} (C_{ik}^{-1} \odot B_{jk}))^{-1}
\end{align}
where the tropical multiplicative inverse is defined as the additive inverse of the regular algebra. The equality holds if and only if element $A_{ij}$ contributions to $C$ (i.e. has nonzero gradient).
Let the mask for $C$ being $\overline C$, the backward rule for gradient masks reads
\begin{align}
\overline{A}_{ij} = \delta(A_{ij}, ((C^{\circ-1} \circ \overline C )B^T)_{ij}^{\circ -1}),
\end{align}
where ${}^{\circ -1}$ is the Hadamard inverse, $\circ$ is the Hadamard multiplication, boolean false is treated as tropical zero and boolean true is treated as tropical one.
This rule defined on matrix multiplication can be easily generalized to the einsum of two tensors by replacing the matrix multiplication between $C^{\circ-1} \circ \overline C$ and $B^T$ by an einsum.

@@

\iffalse
\begin{align}
if |V| = 0 then
return 0
if \exists v \in V with d(v) \leq 1 then
return 1 + mis2(G \backslash N[v])
if \exists v \in V with d(v) = 2 then
(let u 1 and u 2 be the neighbors of v)
if {u 1 , u 2 } \in E then
return 1 + mis2(G \backslash N[v])
if {u 1 , u 2 } \in
/ E then
if |N 2 (v)| = 1 then
(let N 2 (v) = {w})
return max(2 + mis2(G \backslash (N 2 [v] \cup N[w])), 2 + mis2(G \backslash N 2 [v]))
if |N 2 (v)| > 1 then
return max(mis2(G \backslash N[v]), mis2(G \backslash (M(v) \cup cup))
if \exists v \in V with d(v) = 3 then
(let u 1 u 2 and u 3 be the neighbors of v)
if G[N(v)] has no edge then
if v has a mirror then
return max(1 + mis2(G \backslash N[v]), mis2(G \backslash (M(v) \cup {v}))
if v has no mirror then
return max(1 + mis2(G \backslash N[v]), 2 + mis2(G \backslash N[{u 1 , u 2 }]), 2 + mis2(G \backslash
(N[{u_1 , u_3 }] \cup \{u_2\})), 2 + mis2(G \backslash (N[{u_2 , u_3 }] \cup \{1\})))
if G[N(v)] has one or two edges then
return max(1 + mis2(G \backslash N[v]), mis2(G \backslash (M(v) \cup cup))
if G[N(v)] has three edges then
return 1 + mis2(G \backslash N[v])
if \Delta (G) \geq 6 then
choose a vertex v of maximum degree in G
return max(1 + mis2(G \backslash N[v]), mis2(G \backslash v))
if G is disconnected then
(let C \subseteq V be a component of G)
return mis2(G[C]) + mis2(G \backslash C)
if G is 4 or 5-regular then
choose any vertex v of G
return max(1 + mis2(G \backslash N[v]), mis2(G \backslash (M(v) \cup \{v\}))
if \Delta (G) = 5 and \delta (G) = 4 then
choose adjacent vertices v and w with d(v) = 5 and d(w) = 4 in G
return
\max(1 + mis2(G \backslash N[v]), 1 + mis2(G \backslash (\{v\} \cup M(v) \cup N[w])), mis2(G \backslash (M(v) \cup \{v, w\})))
\end{align}
\fi

@@

\centerline{\begin{tikzpicture}
    \def\r{0.1}
    \def\a{0.1}
    \foreach \x in {1,...,5}
        \foreach \y in {1,...,5}
        \node[fill, regular polygon, 
            regular polygon sides=3, scale=0.5] at (\x, \y) {};
    \foreach \x in {1,...,5}
        \foreach \y in {1,...,4}{
            \filldraw[fill=black] (\x-\a,\y+0.5-\a) rectangle (\x+\a, \y+0.5+\a);
            \filldraw[fill=black] (\y+0.5-\a,\x-\a) rectangle (\y+0.5+\a, \x+\a);
            \draw [black,thick] (\x,\y) -- (\x,\y+1);
            \draw [black,thick] (\y,\x) -- (\y+1,\x);
        }
    \foreach \x in {1,...,4}
        \foreach \y in {1,...,4}{
            \filldraw[fill=black] (\x+0.3-\a,\y+0.3-\a) rectangle (\x+0.3+\a, \y+0.3+\a);
            \filldraw[fill=black] (\y+0.3-\a,\x+0.7-\a) rectangle (\y+0.3+\a, \x+0.7+\a);
            \draw [black,thick] (\x,\y) -- (\x+1,\y+1);
            \draw [black,thick] (\y+1,\x) -- (\y,\x+1);
        }
\end{tikzpicture}}

>>>>

One may also want to obtain all ground state configurations, it can be achieved replacing the factors $a_x$ with a set of bit strings $s_x$.
We design a new element type that having algebra
\iffalse
\begin{align}
\begin{split}
    s_x \oplus s_y &= s_x \cup s_y\\
    s_x \odot s_y &= \{\sigma \lor \tau | \sigma \in s_x, \tau \in s_y\}\\
    \mymathbb{0} &= \{\}\\
    \mymathbb{1} &= \{\boldsymbol 0\}
\end{split}
\end{align}
\fi
\begin{align}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases},\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma + \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
    \mymathbb{0} &= \{\}\infty^{-\infty},\\
    \mymathbb{1} &= \{\boldsymbol 0\}\infty^{0},
\end{split}
\end{align}
One can easily check that this replacement does not change the fact that the algebra is a commutative semiring.
We first initialize the bit strings of the variable $x$ in the vertex tensor to a vertex index $i$ dependent onehot vector $x_i = \boldsymbol{e}_{i}$,
then we contract the tensor network. The resulting object will give us the set of all optimal configurations.
By slightly modifying the above algebra, it can also be used to obtain just a single configuration to save computational effort.
We leave this as an exercise for readers.

@@

Moving $B_{ik}$ to the right hand side, we have
\begin{align}
    A_{ij} &\leq (\oplus_{k} (C_{ik}^{-1} \odot B_{jk}))^{-1}
\end{align}
where the tropical multiplicative inverse is defined as the additive inverse of the regular algebra. The equality holds if and only if element $A_{ij}$ contributions to $C$ (i.e. has nonzero gradient).
Let the mask for $C$ being $\overline C$, the backward rule for gradient masks reads
\begin{align}
\overline{A}_{ij} = \delta(A_{ij}, ((C^{\circ-1} \circ \overline C )B^T)_{ij}^{\circ -1}),
\end{align}
where ${}^{\circ -1}$ is the Hadamard inverse, $\circ$ is the Hadamard product, boolean false is treated as tropical zero and boolean true is treated as tropical one.
This rule defined on matrix multiplication can be easily generalized to the einsum of two tensors by replacing the matrix multiplication between $C^{\circ-1} \circ \overline C$ and $B^T$ by an einsum.

@@

@@

\centerline{\begin{tikzpicture}
    \def\r{0.1}
    \def\a{0.1}
    \foreach \x in {1,...,5}
        \foreach \y in {1,...,5}{
        \node[fill, regular polygon, 
            regular polygon sides=3, scale=0.3, rotate=90] at (\x+0.1, \y+0.2) {};
        \draw [black,thick] (\x,\y) -- (\x+0.1,\y+0.2);
        }
    \foreach \x in {1,...,5}
        \foreach \y in {1,...,4}{
            \filldraw[fill=black] (\x-\a,\y+0.5-\a) rectangle (\x+\a, \y+0.5+\a);
            \filldraw[fill=black] (\y+0.5-\a,\x-\a) rectangle (\y+0.5+\a, \x+\a);
            \draw [black,thick] (\x,\y) -- (\x,\y+1);
            \draw [black,thick] (\y,\x) -- (\y+1,\x);
        }
    \foreach \x in {1,...,4}
        \foreach \y in {1,...,4}{
            \filldraw[fill=black] (\x+0.3-\a,\y+0.3-\a) rectangle (\x+0.3+\a, \y+0.3+\a);
            \filldraw[fill=black] (\y+0.3-\a,\x+0.7-\a) rectangle (\y+0.3+\a, \x+0.7+\a);
            \draw [black,thick] (\x,\y) -- (\x+1,\y+1);
            \draw [black,thick] (\y+1,\x) -- (\y,\x+1);
        }
\end{tikzpicture}}
